---
title: "LinkPrediction"
author: "Gregorio Alanis-Lobato"
date: false
output: rmarkdown::github_document
---

Implementation of different classes link predictors and methods to assess and 
visualise their performance.

# Introduction

Complex systems can be represented by graphs $G(V, E)$, where $E$ is the set of
interactions (edges) between a set $V$ of system components (nodes). Link 
predictors assign likelihood scores of interaction to all node pairs that are
disconnected in the observable network topology. These scores can lead to the
prediction of future friendships in a social network or future direct flights 
between cities in an airport transportation network.

To assess the performance of link predictors, one normally removes an increasing
number of edges $E^R \subset E$ from the network of interest, applies a link 
prediction method to the pruned network and checks whether the $|E^R|$ candidate
links with highest likelihood scores are in $E^R$.

`LinkPrediction` implements the following classes of link prediction techniques:

- Neighbourhood-based predictors, which assign high likelihood scores to node 
pairs that share many neighbours (`lp_cn`, `lp_pa`, `lp_jc`, `lp_dice`, `lp_aa`
and `lp_ra`).
- CAR-based predictors, which assign high likelihood scores to node pairs that 
share many neighbours that also interact with themselves (`lp_car`, `lp_cpa`, 
`lp_caa` and `lp_cra`).
- Embedding-based methods rank non-adjacent nodes through distances on a network 
projection in two-dimensional Euclidean space (`lp_isomap`, `lp_leig`, and 
`lp_mce`).
- The Hierarchical Random Graph method, which searches the space of all possible
dendrograms of a network for the ones that best fit its hierarchical structure. 
Non-adjacent node pairs that have high average probability of being connected 
within these dendrograms are considered good candidates for interaction 
(`lp_hrg`).
- The Structural Perturbation Method, which is based on the hypothesis that 
links are predictable if removing them has only small effects on network 
structure (`lp_spm`).

In addition, `LinkPrediction` includes function `lp_matrix` and `get_non_edges`,
which are key to extend the package with other link prediction approaches (more 
on this below).

Finally, the performance of these link predictors on a network of interest can
be assessed and visualised with functions `prune_recover` and 
`plot_lp_precision`.

For more information on the link prediction problem, see:

- Martí́nez, V., Berzal, F. & Cubero, J.-C. A survey of link prediction in
complex networks. *ACM Computing Surveys* **49**, 1-33 (2016).
- Lü̈, L., Pan, L., Zhou, T., Zhang, Y.-C. & Stanley, H. E. Toward link
predictability of complex networks. *PNAS* **112**, 2325–2330 (2015).

# Installation

1. Install the `devtools` package from CRAN if you haven't done so:

```r
install.packages("devtools")
```

2. Load the `devtools` package:

```r
library("devtools")
```

3. Install `LinkPrediction` using the `install_github` function:

```r
install_github("galanisl/LinkPrediction")
```

# Usage

To start using `LinkPrediction`, load the package:

```{r message=FALSE, warning=FALSE}
library("LinkPrediction")
```

`LinkPrediction` includes two complex network datasets (type `?karate_club` and
`?jazz_collab` in R for more information). Let's use the Jazz Collaboration 
network to illustrate the use of the package.

First, let's apply the Common Neighbours link predictor to the network:

```{r message=FALSE, warning=FALSE}
(cn <- lp_cn(jazz_collab))
```

Note how the lp_* functions return a `tibble` of candidate links with their 
likelihoods of interaction. Let's now compute the structural consistency 
$\sigma_c \in [0, 1]$ of this network. The higher it is, the higher its link 
predictability:

```{r}
sigma_c <- structural_consistency(jazz_collab)
mean(sigma_c)
sd(sigma_c)
```

Since the $\sigma_c$ is high, link predictors are likely to give us good 
candidates of interaction.

Let's now implement a link predictor of our own with the help of function 
`get_non_edges`. This function will return a tibble with a random ordering of 
the disconnected node pairs in the network:

```{r}
lp_rnd <- function(g){
  non_edges <- get_non_edges(g)
  prediction <- tibble(nodeA = non_edges[, 1], nodeB = non_edges[, 2],
                       scr = sample(nrow(non_edges))) %>% 
    arrange(desc(scr))
  return(prediction)
}
```

Finally, let's compare our method with other link predictors and visualise the
results:

```{r}
assessment <- prune_recover(jazz_collab, 
                            "lp_rnd", 
                            "lp_cn", 
                            "lp_car", 
                            "lp_isomap")

plot_lp_precision_sd(assessment)
```
